---
title: "DATA 605 Homework 5"
subtitle: "ASSIGNMENT 5"
author: "Stephen Haslett"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
library(ggplot2)
```

#### Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.

The first step to solving this problem is to define **B** and **C**. To do this, we can use R's
[runif()](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/Uniform) function. The **_runif()_** function takes 3 arguments:
**_n_** (_the number of observations we wish to make_), **_min_** (_the lower limit of the distribution_), and finally **_max_**, (_the upper limit of the distribution_).
These arguments combined make this function is a good fit for defining the numbers **B** and **C** as per the assignment instructions.

##### Prove that B and C are proper probability distributions.

```{r DefineNumbers}
# Randomly define numbers B and C from the interval [0, 1], with 10000 observations.
B <- runif(10000, 0, 1)
C <- runif(10000, 0, 1)
```

#### Find the probability that:

##### (a) B + C < 1/2.

```{r SolutionA}

probability_a <- round(sum((B + C) < .5) / 10000, 2)
probability_a
```

**Answer:** _The probability that B + C < 1/2 is **0.13**_.


##### (b) BC < 1/2.

```{r SolutionB}
probability_b <- round(sum((B * C) < .5) / 10000, 2)
probability_b
```

**Answer:** _The probability that BC < 1/2 is **0.84**_.


##### (c) |B − C| < 1/2.

```{r SolutionC}
probability_c <- round(sum(abs((B - C)) < .5) / 10000, 2)
probability_c
```

**Answer:** _The probability that |B − C| < 1/2 is **0.75**_.


##### (d) max{B,C} < 1/2.

```{r SolutionD}
probability_d <- round(sum(pmax(B, C) < .5) / 10000, 2)
probability_d
```

**Answer:** _The probability that max{B,C} < 1/2 is **0.25**_.


##### (e) min{B,C} < 1/2.

```{r SolutionE}
probability_e <- round(sum(pmin(B, C) < .5) / 10000, 2)
probability_e
```

**Answer:** _The probability that min{B,C} < 1/2 is **0.75**_.