---
title: "DATA 605 Discussion 10"
subtitle: "Exercise 7, Chapter 11, Page 414"
author: "Stephen Haslett"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(expm)
```

### Exercise 7
Find the matrices $P^2, P^3, P^4$, and $P^n$ for the Markov chain determined by the transition matrix $P = \begin{pmatrix}1 & 0\ \\0 & 1 \end{pmatrix}$.

Do the same for the transition matrix $P =  \begin{pmatrix}0 & 1\ \\1 & 0 \end{pmatrix}$. Interpret what happens in each of these processes.

### Exercise 7 Solution

```{r exercise7Solution}
# Define the first transition matrix in the problem.
transition_matrix_one <- matrix(c(1, 0, 0, 1), 2)

# Define the second transition matrix in the problem.
transition_matrix_two <- matrix(c(0, 1, 1, 0), 2)
```


#### Find matrices $P^2$ through $P^n$ for the first transition matrix.


**Find the matrix $P^2$**.
```{r firstMatrix}
P2 <- transition_matrix_one%^%2
P2
```


**Find the matrix $P^3$**.
```{r secondMatrix}
P3 <- transition_matrix_one%^%3
P3
```


**Find the matrix $P^4$**.
```{r thirdMatrix}
P4 <- transition_matrix_one%^%4
P4
```


**Find the matrix $P^n$**.
```{r fourthMatrix}
Pn <- transition_matrix_one%^%5
Pn
```


#### Find matrices $P^2$ through $P^n$ for the second transition matrix.

Find the matrix $P^2$.
```{r setTwoFirstMatrix}
P2 <- transition_matrix_two%^%2
P2
```


Find the matrix $P^3$.
```{r setTwoSecondMatrix}
P3 <- transition_matrix_two%^%3
P3
```


**Find the matrix $P^4$.**
```{r setTwoThirdMatrix}
P4 <- transition_matrix_two%^%4
P4
```


Find the matrix $P^n$.
```{r setTwoFourthMatrix}
Pn <- transition_matrix_two%^%5
Pn
```


#### Observations
