---
title: "DATA 605 Homework 2"
subtitle: "ASSIGNMENT 2"
author: "Stephen Haslett"
date: "9/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

## Problem Set 1

#### (1) Show that $A^{T} A \ne AA^{T}$ in general. (Proof and demonstration.)

```{r solution1}

```

#### (2) For a special type of square matrix $A$, we get $A^{T} A = AA^{T}$. Under what conditions could this be true? (Hint: The Identity matrix $I$ is an example of such a matrix).

```{r solution2}

```

## Problem Set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track fights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.

#### Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.



