---
title: "DATA 605 Homework 11"
subtitle: "Assignment 11"
author: "Stephen Haslett"
date: "11/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the necessary libraries.
library(ggplot2)
library(dplyr)
```

### Problem
Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

### Data Analysis

Load the cars dataset into R and take a quick look at the data.
```{r dataSetUp}
# Load the Cars dataset into R.
cars_dataset <- datasets::cars
# Summarize of the data.
str(cars_dataset)
``` 


As we can see above, the Cars dataset comprises of 50 observations of 2 variables - **speed** (_the independent variable_), and **stopping distance (dist)** (_the dependent variable_). 
Each row is an observation of the relation between car speed and the distance it takes for a car to stop.


```{r, dataSummary}
summary(cars_dataset)
```


### Is the relationship between the indepentent and dependant variable linear?

```{r, dataScatter}
plot(dist ~ speed, data = cars_dataset, col = 'blue',
     pch = 16, xlab = 'Speed (Independent Variable)', ylab = 'Stopping Distance (Dependent Variable)',
     main = 'Relationship Between Speed and Stopping Distance')
abline(lm(dist ~ speed, data = cars_dataset), col = 'red')
```

From the above scatter plot we can see that Distance and Speed are positively correlated and whilst not perfectly linear, The relationship appears to be linear.

### Linear Regression Analysis

```{r, dataLinearRegression}
cars_dataset_lm <- lm(speed ~ dist, data = cars_dataset)
summary(cars_dataset_lm)
```

\ 

#### Linear Model Summary Analysis

1. The Multiple R-squared value above is a measurement of how well the model describes the data. For this model, the R-Squared value is **0.6511** which means
that the model explains 65.11% of the data’s variation.

2. The probability value (p-value) that the speed coefficient is not relevant in the model is **1.49e-12** which means that speed is highly relevant in modeling stopping distance. 

3. For well fitted linear regression models, the residuals should be normally distributed with a mean as close to 0 as possible. The residuals median for this model is 0.3615 which is good.

\ 

#### Residual Analysis

```{r, dataParCheck}
par(mfrow = c(2, 2))
plot(cars_dataset_lm)
par(mfrow = c(1, 1))
```

From the **Residuals vs Fitted** plot above we can see that the residuals are not uniformly scattered above and below zero. It appears that as the fitted values increase, so do the residuals.

The **Normal Q-Q** plot above tells us that the residuals from the model are not normally distributed. The 2 ends diverge from the diagonal line indicating that the residuals are not normally distributed.


#### Conclusion
From the above analysis, we can conclude that speed alone is an insufficent predictor in the model to explain the data and to predict stopping distance. To get a more accurate picture, we would most likely need to introduce additional predictors to our model such as weather, car weight, effectiveness of the car's brake pads, etc.