---
title: "DATA 605 Homework 5"
subtitle: "ASSIGNMENT 5"
author: "Stephen Haslett"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
library(ggplot2)
```

#### Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.

The first step to solving this problem is to define **B** and **C**. To do this, we can use R's
[runif()](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/Uniform) function. The **_runif()_** function takes 3 arguments:
**_n_** (_the number of observations we wish to make_), **_min_** (_the lower limit of the distribution_), and finally **_max_**, (_the upper limit of the distribution_).
These arguments combined make this function is a good fit for defining the numbers **B** and **C** as per the assignment instructions.

```{r defineBandC}
# Randomly define B and C from the interval [0, 1].
B <- runif(100, 0, 1)
C <- runif(100, 0, 1)
B
C
```

#### Find the probability that:

##### (a) B + C < 1/2.

```{r solutionA}

```

**Answer:** _ _


##### (b) BC < 1/2.

```{r solutionB}

```

**Answer:** _ _


##### (c) |B − C| < 1/2.

```{r solutionC}

```

**Answer:** _ _


##### (d) max{B,C} < 1/2.

```{r solutionD}

```

**Answer:** _ _


##### (e) min{B,C} < 1/2.

```{r solutionE}

```

**Answer:** _ _






