---
title: "DATA 605 Homework 7"
subtitle: "ASSIGNMENT 7"
author: "Stephen Haslett"
date: "10/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(Matrix)
```

**1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of**
**which is uniformly distributed on the integers from 1 to k. Let Y denote the**
**minimum of the Xi’s. Find the distribution of Y.**


```{r Answer1}

```

**2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has **
**a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. **
**(Include the probability statements and R Code for each part.).**


**2. a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric.**
**(Hint: the probability is equivalent to not failing during the first 8 years..)**

```{r Answer2-a}
```

**2. b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.**

```{r Answer2-b}
```


**2. c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial.**
**(Hint: 0 success in 8 years)**

```{r Answer2-c}
```

**2. d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.**

```{r Answer2-c}
```

