---
title: "DATA 605 Homework 8"
subtitle: "ASSIGNMENT 8"
author: "Stephen Haslett"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(Matrix)
```


#### Exercise 11 - Page 303
A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?


```{r Exercise-11-solution}
# Sum of the bulbs.
number_of_bulbs <- 100

# The mean lifetime of the bulbs is 1000 hours.
bulbs_mean_lifetime <- 1000

# We have 100 bulbs, so to find the lifetime of the first lightbulb (minimum lifetime),
# we divid the mean lifetime by the number of bulbs.
first_bulp_lifespan <- 1000/100

# The first lightbulb will burn out after 10 hours.
first_bulp_lifespan
```

**Answer:** The expected time for the first of these bulbs to burn out is **10 hours**.


#### Exercise 14 - Page 303
Assume that  $x_{1}$ and $x_{2}$ are independent random variables, each having an exponential density with parameter $\lambda$. Show that $Z = x_{1} − x_{2}$ has density:
$$f_{Z} = (1/2)\lambda_{e}^-\lambda|z|$$.


```{r Exercise-14-solution}

```

#### Exercise 1 - Pages 320 - 321
Let $X$ be a continuous random variable with mean $µ = 10$ and variance $σ^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.

(a) $P(|X − 10| ≥ 2)$.
(b) $P(|X − 10| ≥ 5)$.
(c) $P(|X − 10| ≥ 9)$.
(d) $P(|X − 10| ≥ 20)$.

  
```{r Exercise-1-solution}

```