---
title: "DATA 605 Homework 13"
subtitle: "Assignment 13"
author: "Stephen Haslett"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the necessary libraries.
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
```

#### 1. Use integration by substitution to solve the integral below:

$\int 4e^{-7x}dx$

\ 

##### 1. Solution:
Let $u = -7x$

$du = -7 dx$

$dx = \frac{du}{-7}$

$\frac{4}{-7}\int e^u du$

$\frac{-4}{7}e^{-7x} + C$

\ 

#### 2. Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{\text{d}N}{\text{d}t} = -\frac{3150}{t^4} - 220$ bacteria per cubic centimeter per day, where $t$ is the number of days since treatment began. Find a function $N(t)$ to estimate the level of contamination if the level after day 1 was 6530 per cubic centimeter.

\ 

##### 2. Solution:


$\frac{\text{d}N}{\text{d}t} = -\frac{3150}{t^4} - 220 \ \to \ \text{d}N = (-\frac{3150}{t^4}-220)\text{d}t$

Find $N$

$N = \int (-\frac{3150}{t^4}-220)\text{d}t \ \ = \int -3150(t^{-4}) \text{d}t - \int 220\text{d}t$


Apply the power rule:

$N = \frac{-3150}{-3}(t^{-3}) - 220t + C$.

\ 

Solve for $N(1) = 6530$:

$N(1) \frac{-3150}{-3}(1^{-3}) - 220(1) + C = 6530$

$C = 6530 - 1050 + 220 = 5700$.

$N(t) = -1050(t^{-3}) - 200(t) + 5700$.






```{r}
```

