---
title: "DATA 605 Homework 10"
author: "Stephen Haslett"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem
Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if:


(a) He bets 1 dollar each time (timid strategy).

(b) He bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

(c) Which strategy gives Smith the better chance of getting out of jail?


### Solutions

Define the universal variables used in the game.
```{r UniversalVariables}
# Define the probability of winning.
p <- 0.4

# Define the probability of losing.
q <- 0.6

# Calculate the ratio of winning probability to losing probability.
ratio <- q/p

# Define the minimum bet amount, and target win amount.
bet <- 1
win_target <- 8
```


**(a) Find the probability that he wins 8 dollars before losing all of his money if he bets 1 dollar each time (timid strategy)**.

```{r SolutionA}
# Calculate the probability of the prisoner winning the game.
win_probability <- (1 - ratio^bet) / (1 - ratio^win_target)
probability_prisoner_wins <- round(win_probability, 2)
probability_prisoner_wins
```

**Answer:** The probability that the prisoner will win the game is **0.02**.


