---
title: "DATA 605 Homework 6"
subtitle: "ASSIGNMENT 6"
author: "Stephen Haslett"
date: "10/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
library(ggplot2)
```

**1. A box contains 54 red marbles, 9 white marbles, and 75 blue marbles. If a marble is randomly selected from the box,**
**what is the probability that it is red or blue? Express your answer as a fraction or a decimal number rounded to four**
**decimal places.**

The first step to soving the problem us to find out how many marbles we have in total.
```{r AnswerOne}
# We have 54 red marbles, 9 white marbles, and 75 blue marbles.
red_marbles <- 54
white_marbles <- 9
blue_marbles <- 75

# Find the sum of all marbles.
total_marbles <- red_marbles + white_marbles + blue_marbles
total_marbles
```

Now that we know we have a total of **138** marbles, what is the probability that the marble we pick from the box is red?
Similarly, what is the probability that the marble we pick from the box is blue?

- There are **54 red marbles**, so the probability of picking a red marble from the box is: $\frac{54}{138}$

- There are **75 blue marbles**, so the probability of picking a blue marble from the box is: $\frac{75}{138}$

