---
title: "DATA 605 Homework 4"
subtitle: "ASSIGNMENT 4"
author: "Stephen Haslett"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(Matrix)
```

## Problem Set 1

#### Verify using R that SVD and Eigenvalues are related.

**A. Given a 3 X 2 matrix A, write code in R to compute $X = AA^T$ and $Y = A^TA$.**
$$A = \begin{bmatrix}1 & 2 & 3 \\ -1 & 0 & 4\end{bmatrix}$$

```{r Set-1A}
# Define matrix A.
A <- matrix(c(1, 2, 3, -1, 0, 4), 2, 3)

# Calculate X and Y.
X <- A %*% t(A)
Y <- t(A) %*% A
```

**B. Compute the eigenvalues and eigenvectors of X and Y using the built-in commands in R.**

```{r Set-1B}
# Compute the eigenvalues and eigenvectors of X and Y.
eigens_x <- eigen(X)
eigens_y <- eigen(Y)
```

**_Eigenvalues of X._**

```{r Set-1C-x-eigenvalues}
eigens_x$values
```

**_Eigenvectors of X._**

```{r Set-1C-x-eigenvectors}
eigens_x$vectors
```

**_Eigenvalues of Y._**

```{r Set-1C-y-eigenvalues}
eigens_y$values
```

**_Eigenvectors of Y._**

```{r Set-1C-y-eigenvectors}
eigens_y$vectors
```

**C. Compute the left-singular, singular values, and right-singular vectors of A using the svd command.**

```{r Set-1C}
# Compute the left-singular vector of A.
left_singular_vectors <- svd(A)$u

# Compute the singular values of A.
singular_values <- svd(A)$d

# Compute the right-singular vector of A.
right_singular_vectors <- svd(A)$v
```

**_Left singular vectors of A._**

```{r Set-1C-left-singular-vectors}
left_singular_vectors
```

**_Singular values of A._**

```{r Set-1C-singular-values}
singular_values
```

**_Right singular vectors of A._**

```{r Set-1C-right-singular-vectors}
right_singular_vectors
```

**D. Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y.**

```{r Set-1D-examination-x}
# Compare the left singular vectors of A with the eigenvectors of X.
left_singular_vectors
eigens_x$vectors

# Confirm that they are indeed equal.
round(abs(left_singular_vectors)) == round(abs(eigens_x$vectors))
```

```{r Set-1D-examination-y}
# Compare the right singular vectors of A with the eigenvectors of Y.
right_singular_vectors
eigens_y$vectors
```

**E. Show that the two non-zero eigenvalues of both X and Y are the same.**

```{r Set-1E-value-comparision}
# Compare the two non-zero eigenvalues of both X and Y to confirm that they the same.
eigens_y$values
eigens_x$values
```

**F. Show that the two non-zero eigenvalues of both X and Y are squares of the non-zero singular values of A.**
```{r Set-1F-square}
# Compare the two non-zero eigenvalues of both X and Y to confirm that they the same.
eigens_x$values

# Eigenvalues of Y.
eigens_y$values

# Singular values of A.
singular_values^2

# Confirm that the eigenvalues of X are squares of the non-zero singular values of A.
round(abs(eigens_x$values[1:2])) == round(abs(singular_values^2))

# Confirm that the eigenvalues of Y are squares of the non-zero singular values of A.
round(abs(eigens_y$values[1:2])) == round(abs(singular_values^2))
```


## Problem Set 2

#### Write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors.

Your function should have the following signature: B = myinverse(A) where A is a matrix and B is its inverse and AxB = I. The off-diagonal
elements of I should be close to zero, if not zero. Likewise, the diagonal elements should be close to 1, if not 1.

```{r Set-2-inverse-function}
# Define the function myinverse().
myinverse <- function(A) {
  # Check if the passed matrix is a square matrix.
  if (nrow(A) != ncol(A)) {
    return('The provided matrix is not a square matrix.')
  }

  # Check to ensure that the passed matrix is invertable.
  if (det(A) == 0) {
    return('The provided matrix is not invertable.')
  }

  # Assign the matrix rows and columns to variables (rows and columns).
  rows <- nrow(A)
  columns <- ncol(A)

  # Define the co-factor matrix.
  cofactor_matrix <- diag(0, rows, columns)
  for(i in 1:rows) {
    for(j in 1:columns) {
      # Populate the matrix with rows and columns.
      cofactor_matrix[i,j] <- (-1) ^ (i + j) * det(A[-i, -j])
    }
  }

  # Define the inverted matrix and return it.
  inverted_matrix <- (t(cofactor_matrix) / det(A))

  return(inverted_matrix)
}
```

**Test the myinverse() function with a 3x3 matrix.**
```{r Set-2-inverse-function-test-3x3}
# Define the 3x3 matrix A.
A <- matrix(c(1, 0, 5, 3, 1, 5, 3, 4, 0), 3, 3)
A

# Pass the matrix to the myinverse() function and print the inverted matrix.
B <- myinverse(A)
B

# Check that the result is correct.
accuracy_check <- solve(A)
round(accuracy_check, 2) == round(B, 2)

# Check that the off-diagonal elements of I are zero, or close to zero,
# and that the diagonal elements are 1, or close to 1.
round(A %*% B, 2)
```

**Test the myinverse() function with a 4x4 matrix.**
```{r Set-2-inverse-function-test-4x4}
# Define the 4x4 matrix A.
A <- matrix(c(1, 6,-2, -3, 2, -8, 3, -6, -1, 0, 3, 16, -5, -2, 7, 9), 4, 4)
A

# Pass the matrix to the myinverse() function and print the inverted matrix.
B <- myinverse(A)
B

# Check that the result is correct.
accuracy_check <- solve(A)
round(accuracy_check, 2) == round(B, 2)

# Check that the off-diagonal elements of I are zero, or close to zero,
# and that the diagonal elements are 1, or close to 1.
round(A %*% B, 2)
```
