---
title: "DATA 605 Homework 15"
author: "Stephen Haslett"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\ 

##### 1. Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

$$( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )$$

```{r, Answer1}
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

regression <- data.frame (x,y)

# Round to the nearest hundredth.
round(coef(lm(y ~ x, regression)), 2)
```

**Answer:** The equation of the regression line is $y = -14.8 + 4.26x$

\ 

##### 2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form $( x, y, z )$. Separate multiple points with a comma.

$$f ( x, y ) = 24x - 6xy^2 - 8y^3$$


**Answer:**

Partial derivatives:

- $f_x = 24 - 6y^2$

- $f_y = -12xy -24y^2$

Set $f_x$ to 0 and solve for $y$:

- $24 - 6_y^2 = 0$

- $24 = 6_y^2$

- $4 = y^2$

- $±2 = y$

Set $f_y$ to 0 and substitute $y = 2$:

- $−12xy − 24y^2 = 0$

- $xy + 2y^2 = 0$

- $x(2) + 2(2)^2 = 0$

- $2_x + 8 = 0$

- $2X = -8$

- $x = -4$

Substitute $y = -2$:

- $x(-2) + 2(-2)^2 = 0$

- $-2x + 8 = 0$

- $−2x = −8$

- $x = 4$


Substitute $(x,y)$ for $z$:

- $f(x,y) = 24x - 6xy^2 - 8y^3$

Substitute $(-4,2)$ and $(4,-2)$:

- $z = 24(-4) - 6(-4)(2)^2 - 8(2)^3$

- $z = -96 + 96 - 64$

- $z = -64$

- $z = 24(4) - 6(4)(-2)^2 - 8(-2)^3$

- $z = 96 - 96 + 64$

- $z = 64$

**The critical points are $(-4, 2, -64)$ and $(4, -2, 64)$**

Now use the Second Derivative test to determine if the points are minimum, maximum, or saddle.

- $f_{xx} = 0$

- $f_{yy} = -12x - 48y$

- $f_{xy} = - 12y$

- $D = f_{xx}(x,y) * f_{yy}(x,y) - f_{xy}^2(x,y)$

- $D = 0(x,y) * (-12x - 48y(x,y)) - (- 12y)^2(x,y)$

- $D = 0 - 144y^2(x,y)$

- $D = -144y^2(x,y)$

**Answer:** As per the Second Derivative Test, $D(x,y) < 0$ for both $(-4,2,-64)$ and $(4,-2,64)$, so therefore both critical points are saddle points.

\ 

##### 3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for $x$ dollars and the "name" brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

**Step 1.** Find the revenue function $R ( x, y )$.

**Answer**:

The revenue function is unit cost multiplied by units sold for the house/name brand and can be defined as follows:

$$R(x,y) = x(81 - 21x + 17y) + y(40 + 11x -23y)$$


**Step 2.** What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?

**Answer**:

We can find the revenue by substituting $x$ in the above function with the price of the house brand (**$2.30**), and $y$ with the price of the name brand (**$4.10**).

```{r, Step2Answer}
x <- 2.3
y <- 4.1

revenue <- (x * (81 - (21 * x) + (17 * y))) + (y * (40 + (11 * x) - (23 * y)))
revenue
```


**The revenue is $116.62.**

\ 

##### 4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$, where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

If $x + y = 96$, then $y = 96 - x$.

$$C(x,y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$$
$$C(x,y) = \frac{1}{6}x^2 + \frac{1}{6}(96-x)^2 + 7x + 25(96-x) + 700$$
$$C(x,y) = \frac{1}{6}x^2 + \frac{1}{6}(9216 - 192x + x^2) + 7x + 2400 - 25x + 700$$
$$C(x,y) = \frac{1}{6}x^2 + 1536 - 32x + \frac{x^2}{6} + 7x + 2400 - 25x + 700$$
$$C(x,y) = \frac{1}{3}x^2 - 50x + 4636$$
$$C' = \frac{2}{3}x - 50$$
$$0 = \frac{2}{3}x - 50$$
$$50 = \frac{2}{3}x$$
$$75 = x$$

**Answer**: 75 units should be produced in Los Angeles and the remaining 21 units should be produced in Denver to minimize the total weekly cost.

\ 

##### 5. Evaluate the double integral on the given region.

$$\iint_{R}(e^8x + 3y) \ dA; R: 2 \leq x \leq 4 \ and \ 2 \leq y \leq 4$$

**Answer**:

$$\int_2^4\int_2^4(e^{8x + 3y})dx dy$$
$$int_2^4(1/8)(e^{16} - 1)e^{3y + 16} dy$$
$$e^{16} - 1) / 24 e^{3y + 16} + C$$
$$(e^{44} - e^{28}) / 24 - (e^{38} - e^{22}) / 24$$
$$(1 / 24)(e^{22} - e^{28} - e^{38} + e^{44})$$