---
title: "DATA 605 Homework 4"
subtitle: "ASSIGNMENT 4"
author: "Stephen Haslett"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(Matrix)
```

## Problem Set 1

#### Verify using R that SVD and Eigenvalues are related.

**A. Given a 3 X 2 matrix A, write code in R to compute $X = AA^T$ and $Y = A^TA$.**
$$A = \begin{bmatrix}1 & 2 & 3 \\ -1 & 0 & 4\end{bmatrix}$$

```{r Set-1A}
# Define matrix A.
A <- matrix(c(1, 2, 3, -1, 0, 4), 2, 3)

# Calculate X and Y.
X <- A %*% t(A)
Y <- t(A) %*% A
```

**B. Compute the eigenvalues and eigenvectors of X and Y using the built-in commands in R.**

```{r Set-1B}
# Compute the eigenvalues and eigenvectors of X and Y.
eigens_x <- eigen(X)
eigens_y <- eigen(Y)
```

**_Eigenvalues of X._**

```{r Set-1C-x-eigenvalues}
eigens_x$values
```

**_Eigenvectors of X._**

```{r Set-1C-x-eigenvectors}
eigens_x$vectors
```

**_Eigenvalues of Y._**

```{r Set-1C-y-eigenvalues}
eigens_y$values
```

**_Eigenvectors of Y._**

```{r Set-1C-y-eigenvectors}
eigens_y$vectors
```

**C. Compute the left-singular, singular values, and right-singular vectors of A using the svd command.**

```{r Set-1C}
# Compute the left-singular vector of A.
left_singular_vectors <- svd(A)$u

# Compute the singular values of A.
singular_values <- svd(A)$d

# Compute the right-singular vector of A.
right_singular_vectors <- svd(A)$v
```

**_Left singular vectors of A._**

```{r Set-1C-left-singular-vectors}
left_singular_vectors
```

**_Singular values of A._**

```{r Set-1C-singular-values}
singular_values
```

**_Right singular vectors of A._**

```{r Set-1C-right-singular-vectors}
right_singular_vectors
```

**D. Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y.**

```{r Set-1D-examination-x}
# Compare the left singular vectors of A with the eigenvectors of X.
left_singular_vectors
eigens_x$vectors

# Confirm that they are indeed equal.
round(abs(left_singular_vectors)) == round(abs(eigens_x$vectors))
```

```{r Set-1D-examination-y}
# Compare the right singular vectors of A with the eigenvectors of Y.
right_singular_vectors
eigens_y$vectors
```

**E. Show that the two non-zero eigenvalues of both X and Y are the same.**

```{r Set-1E-value-comparision}
# Compare the two non-zero eigenvalues of both X and Y to confirm that they the same.
eigens_y$values
eigens_x$values
```

**F. Show that the two non-zero eigenvalues of both X and Y are squares of the non-zero singular values of A.**
```{r Set-1F-square}
# Compare the two non-zero eigenvalues of both X and Y to confirm that they the same.
eigens_x$values

# Eigenvalues of Y.
eigens_y$values

# Singular values of A.
singular_values^2

# Confirm that the eigenvalues of X are squares of the non-zero singular values of A.
round(abs(eigens_x$values[1:2])) == round(abs(singular_values^2))

# Confirm that the eigenvalues of Y are squares of the non-zero singular values of A.
round(abs(eigens_y$values[1:2])) == round(abs(singular_values^2))
```
