---
title: "DATA 605 Discussion 12"
subtitle: "Is height hereditary?"
author: "Stephen Haslett"
date: "11/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the necessary libraries.
library(tidyverse)
library(knitr)
library(HistData)
```

**Using simple linear regression, we want to answer the question "Is there a relationship between a son's height and his father's height?"**

### The Dataset
For this investigation we will use a dataset compiled by [Francis Galton](https://en.wikipedia.org/wiki/Francis_Galton) that contains data about the heights of children and their parents.

First we will import the dataset into R (the dataset can be accessed via the HistData R package) and take a brief look at the data.
```{r dataSetUp}
# Import Galton's dataset from the HistData package.
data('GaltonFamilies')
galton_height_data <- GaltonFamilies
kable(head(galton_height_data), format = 'markdown')

# Count the number of observations in the dataset.
nrow(galton_height_data)
```

As we can see from the above, the dataset contains 8 variables: _family_, _father_, _mother_, _midparentHeight_, _children_, _childNum_, _gender_, and _childHeight_. The dataset contains **934** observations. Since we are only interested in how a father's height influences his son's height, we can ignore a lot of these variables and create a new dataset containing data relevant to sons and father.

```{r, fatherSonDataSet}
# Create a new dataset containing information releveant to fathers and their sons.
father_son_heights <- GaltonFamilies %>%
  filter(gender == 'male') %>%
  group_by(family) %>%
  sample_n(1) %>%
  ungroup() %>%
  select(father, childHeight) %>%
  rename(son = childHeight)

kable(head(father_son_heights), format = 'markdown')
```




### Data Exploration
The first step in our data exploration is to take a look at a mean, and standard deviation for the father and son variables.

```{r, dataSummary}
# Summarize the dataset.
options(knitr.kable.NA = '')
father_son_heights %>% summarize(mean(father), sd(father), mean(son), sd(son))
```

Looking at the data summary, 


```{r, medianAssetsScatter}

```




## Regression Model

```{r, dataLinearRegression}

```



```{r, data}
# Ensure R outputs the summary data in a non scientific format so that it is more readable
# (i.e. '-691.837588' rather than '-6.918e+02').
#options(scipen = 4)

# Output a summary of the data.

```


## Residual Analysis

```{r residualAnalysis}

```




