---
title: "DATA 605 Homework 12"
subtitle: "Assignment 12"
author: "Stephen Haslett"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the necessary libraries.
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
```

### Problem
The who.csv dataset contains real-world data from 2008. The variables included follow:

Country: name of the country

LifeExp: average life expectancy for the country in years

InfantSurvival: proportion of those surviving to one year or more

Under5Survival: proportion of those surviving to five years or more

TBFree: proportion of the population without TB.

PropMD: proportion of the population who are MDs

PropRN: proportion of the population who are RNs

PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate

GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate

TotExp: sum of personal and government expenditures.


### Load the dataset into R and take a glance at it's structure.
```{r dataSetUp}
# Load the who dataset into R.
who_dataset <- read.csv('https://raw.githubusercontent.com/stephen-haslett/data605/data605-week-12/who.csv')

# Attach the dataset so we can access the variables easily.
attach(who_dataset)

# View the first 10 observations in the dataset to get a sense of how the data is structured.
kable(head(who_dataset, 10),  format = 'markdown')
``` 

\ 

### Questions

1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

\ 

##### Scatterplot of LifeExp~TotExp

```{r answerOne}
# Create a scatter plot of life expectancy by mean government expenditures per capita on healthcare.
life_expectancy_plot <- ggplot(who_dataset, aes(x = TotExp, y = LifeExp)) + geom_point(color = "salmon") +
    ylab("Life Expectancy") + xlab("Government Expenditures Per Capita on Healthcare")

life_expectancy_plot
```


##### Provide and interpret the F statistics, R^2, standard error, and p-values

```{r answerOneRegression}
# Simple linear regression.
who_simple_regression <- lm(LifeExp ~ TotExp, who_dataset)

# Run the summary function on the model so we can interpret the F statistics, 
# R-squared, standard error, and p-values.
summary(who_simple_regression)
```

**F-statistic:** The F-statistic is high enough for us to reject the null hyopthesis that "_government expenditures on healthcare do not contribute to a country’s life expentancy_". We can therefore say that the "TotExp" variable does have an effect on the "LifeExp" variable.

**R-squared:** The R-squared value is low (0.2577), suggesting that the model only explains 25.77% of the data variation.

**Standard error:** The standard error is significantly high which suggests that the model does not fit well. 

**P-values:** The low p-values suggest that govenerment expenditure does contribute to a country's life expectancy. Therefore the 'TotExp'variable does contribute to the model. 

\ 


2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"

\ 

##### Raise life expectancy to the 4.6 power (LifeExp^4.6). Raise total expenditures to the 0.06 power (TotExp^.06)

```{r answerTwoRaise}
# Raise the life expectancy variable to the power of 4.6.
life_expectancy_46 = who_dataset$LifeExp^(4.6)

# Raise the total expenditures variable to the power of 0.6.
total_expenditures_06 = who_dataset$TotExp^(0.06)
```

\ 

##### Plot LifeExp^4.6 as a function of TotExp^.06, and re-run the simple regression model using the transformed variables.

```{r answerTwoPlot}
# Create a scatter plot for LifeExp^4.6 as a function of TotExp^.06.
plot(total_expenditures_06, life_expectancy_46, main='LifeExp^4.6 as a function of TotExp^.06',
     ylab = 'Life Expectancy',
     xlab = 'Total Expenditure',
     col = 2)

# Re-run the simple regression model on the adjusted variables.
adjusted_model = lm(life_expectancy_46 ~ total_expenditures_06)
abline(adjusted_model, col = 1)
```

\ 

##### Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?".

```{r answerTwoSummary}
# Run the summary function on the adjusted  model so we can interpret
# the F statistics, R-squared, standard error, and p-values.
summary(adjusted_model)
```

**F-statistic:** The F-statistic for this model is much higher than that of the unadjusted model in question 1. The p-value is also much lower and therefore we can confirm that this model is much better than the unadjusted model in question 1.

**R-squared:** The R-squared value (0.7298) is also much higher than that of the unadjusted model. The unadjusted model only explained 25.77% of the data variation whereas this model explains 72.98% of the data variation.

**Standard error:** The standard error for this model is surprisingly high considering the positive F-statistic and R-squared values. This may be explained by the fact that we raised the LifeExp and TotExp variables exponentially.

**P-values:** The p-values for this model are much lower than those of the unadjusted model giving us more confidence that the 'TotExp'variable contributes to the model.

\ 

3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.



4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
_LifeExp = b0+b1 x PropMd + b2 x TotExp + b3 x PropMD x TotExp_

