---
title: "DATA 605 Discussion 12"
subtitle: "Is height hereditary?"
author: "Stephen Haslett"
date: "11/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the necessary libraries.
library(dplyr)
library(tidyverse)
library(knitr)
library(HistData)
library(ggplot2)
```

**Using simple linear regression, we want to answer the question "Is there a relationship between a child's height and the height of their parents?"**

### The Dataset
For this investigation we will use a dataset compiled by [Francis Galton](https://en.wikipedia.org/wiki/Francis_Galton) that contains data about the heights of children and their parents.

First we will import the dataset into R (the dataset can be accessed via the HistData R package) and take a brief look at the data.
```{r dataSetUp}
# Import Galton's dataset from the HistData package.
data('Galton')
galton_height_data <- GaltonFamilies
kable(head(galton_height_data), format = 'markdown')

# Count the number of observations in the dataset.
nrow(galton_height_data)
```

As we can see from the above, the dataset contains **934** observations, and 8 variables: _family_, _father_, _mother_, _midparentHeight_, _children_, _childNum_, _gender_, and _childHeight_.

In order to assess if there is a relationship between a child's height and their parents' height, we can plot the data
in a scatter graph. This allows us to easily assess if a relationship exists.

```{r, parentChildHeightScatterPlot}
# PLot a scatter plot of the Galton dataset.
galton_plot <- ggplot(Galton, aes(x = parent, y = child)) + geom_point(color = "blue") +
    ylab("Child's Height in Inches") + xlab("Parents' Height in Inches")

galton_plot
```


### Data Exploration
Looking at the above scatter plot, it does appear that there is a relationship between the height of a child's parents, and the height of the child. As the parents' height increases, so does the child's height. To confirm this, we can take a look at the correlation between the 2 variables (child's height vs parents' height).  

##### Correlation

```{r, correlation}
# Calculate the correlation between the 2 variables.
round(cor(Galton$parent, Galton$child), 2)
```

The correlation between the 2 variables is **0.46** which confirms that there is a strong relationship between the 2 variables.

##### Distribution normality of the dependent variable (child's height)

The next thing we want to check is the distribution normality for the dependent variable (child's height). We can check this by plotting a histogram and seeing
if the distribution is bell shapped.


```{r, distributionNormality}
hist(Galton$child)
```

The distribution is close to normal (bell shapped) and therefore we can proceed to check the linearity of the variables.

##### Linearity of the relationship between parents' height and child's height

```{r linearity}
plot(parent ~ child, Galton)
```

The relationship between a child's height and their parents' height is not overly apparent, but it is linear and thus we can move forward with building the regression model. 


## Regression Model
The goal of our regression model is to predict the height of a child (_response/dependent variable_) based on their parents' height (_predictor/independent variable_).

##### Build the regression model
The first step is to build the regression model using R's lm() function.

```{r, regressionModel}
galton_height_model <- lm(parent ~ child, Galton)
summary(galton_height_model)
```








```{r, regressionPlot}
regression_plot <-ggplot(Galton, aes(x = parent, y = child)) + geom_point(color = "blue") + 
    ylab("Child's height in inches") + xlab("Parents' average height in inches") + 
    geom_smooth(data = Galton, method = "lm", se = FALSE, color = "black", size = 0.5)

regression_plot
```



```{r, data}
# Ensure R outputs the summary data in a non scientific format so that it is more readable
# (i.e. '-691.837588' rather than '-6.918e+02').
#options(scipen = 4)

# Output a summary of the data.

```


## Residual Analysis

```{r residualAnalysis}

```




