---
title: "DATA 605 Homework 10"
author: "Stephen Haslett"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem
Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if:


(a) He bets 1 dollar each time (timid strategy).

(b) He bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

(c) Which strategy gives Smith the better chance of getting out of jail?

### Solutions

Define the universal variables used in the game.
```{r UniversalVariables}
# Define the probability of losing (q).
q <- 0.6

# Define the probability of winning (p).
p <- 0.4

# Calculate the probability ratio of winning to losing (q/p).
ratio <- q/p

# Define the minimum bet amount (x), and the amount needed to win the bet (y).
x <- 1
y <- 8
```


**(a) Find the probability that he wins 8 dollars before losing all of his money if he bets 1 dollar each time (timid strategy)**.

We can calculate the probability of the prisoner winning \$8 before losing all of his money utilizing the following formula:
$$P_{x} = \frac{(1 - \frac{q}{p})^{x}}{(1 - \frac{q}{p})^{y}}$$.

Using the above formula, we can use R to calculate the probability of the prisoner winning as follows:

```{r SolutionA}
# Calculate the probability of the prisoner winning the game (timid strategy).
win_probability <- (1 - ratio^x) / (1 - ratio^y)
probability_prisoner_wins <- round(win_probability, 2)
probability_prisoner_wins
```

**Answer:** The probability that the prisoner will win the game is **0.02**.


**(b) He bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy)**.

```{r SolutionB}
# Calculate the probability of the prisoner winning the game (bold strategy).

```