---
title: "DATA 605 Homework 7"
subtitle: "ASSIGNMENT 7"
author: "Stephen Haslett"
date: "10/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(Matrix)
```

**1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of**
**which is uniformly distributed on the integers from 1 to k. Let Y denote the**
**minimum of the Xi’s. Find the distribution of Y.**

To answer this question, we can plot a histogram and observe Y's distribution.

```{r Answer1}
# Define the variables we need to simulate the problem.
n <- 50000
k <- 50
samples <- 10
Y <- c()

# Simulate the problem.
for (count in 1:n) {
  Xn <- sample(1:k, samples, TRUE)
  Y <- c(Y, min(Xn))
}

# Plot the results as a histogram so we can observe the distribution of Y.
distribution_of_y <- hist(Y, col = '#A52A2A', main = 'Distribution of Y')
```


**Answer:** The distribution of **Y** is positively skewed.

**2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has **
**a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. **
**(Include the probability statements and R Code for each part.).**

```{r Answer2-setup}
  failure_probability <- 0.1
```

**2. a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric.**
**(Hint: the probability is equivalent to not failing during the first 8 years..)**

```{r Answer2-aa}
  # Calculate the probability that the machine will fail after 8 years using a geometric model.
  machine_failure_geo_probability <- pgeom(8, failure_probability, lower.tail = FALSE)
  machine_failure_geo_probability
```

```{r Answer2-ab}
  # Calculate the standard deviation for the geometric model.
  geometric_standard_deviation <- sqrt((1 - failure_probability) / (failure_probability^2))
  geometric_standard_deviation
```

**2. b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.**

```{r Answer2-ba}
  # Calculate the probability that the machine will fail after 8 years using an exponential model.
  machine_failure_exp_probability <- pexp(8, failure_probability, lower.tail = FALSE)
  machine_failure_exp_probability
```

```{r Answer2-bb}
  # Calculate the standard deviation for the exponential model.
  exponential_standard_deviation <- sqrt(1 / failure_probability^2)
  exponential_standard_deviation
```

**2. c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial.**
**(Hint: 0 success in 8 years)**

```{r Answer2-ca}
  # Calculate the probability that the machine will fail after 8 years using an binomial model.
  machine_failure_binomial_probability <- pbinom(0, 8, failure_probability)
  machine_failure_binomial_probability
```


```{r Answer2-cb}
  # Calculate the standard deviation for the binomial model.
  binomial_standard_deviation <- sqrt(8 * failure_probability * (1 - failure_probability))
  binomial_standard_deviation
```


**2. d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.**

```{r Answer2-da}
  # Calculate the probability that the machine will fail after 8 years using an Poisson model.
  machine_failure_poisson_probability <- ppois(0, 0.8)
  machine_failure_poisson_probability
```

```{r Answer2-db}
  # Calculate the standard deviation for the Poisson model.
  poisson_standard_deviation <- sqrt(8 * failure_probability)
  poisson_standard_deviation
```