---
title: "DATA 605 Final Project"
author: "Stephen Haslett"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\ 

#### Problem 1

Using R, generate a random variable $X$ that has 10,000 random uniform numbers from $1$ to $N$, where $N$ can be any number of your choosing greater than or equal to $6$. Then generate a random variable $Y$ that has $10,000$ random normal numbers with a mean of $\mu =\sigma=(N+1)/2$.  

```{r, randomVariableX}
set.seed(100)

# Set N to any number greater than or equal to 6 (in this case, 8).
N <- 8

# Generate a random variable X that has 10,000 random uniform numbers from 1 to N.
X <- runif(10000, 1, N)

# Generate a random variable Y that has 10,000 random normal numbers with the requested mean.
Y <- rnorm(10000, mean = (N+1)/2, (N+1)/2)
```

\ 

##### _Problem 1 - Probability_

Calculate as a minimum the below probabilities $a$ through $c$.  Assume the small letter "$x$" is estimated as the median of the $X$ variable, and the small letter "$y$" is estimated as the 1st quartile of the $Y$ variable. Interpret the meaning of all probabilities.

```{r, probability}
# Small x is estimated as the median of the X variable.
x <- median(X)

# Small y is estimated as the 1st quartile of the Y variable.
y <- quantile(Y, 0.25)
```

A. $P(X>x \ | \ X>y)$

```{r, probabilityA}
segment_one <- (X > x)
probability_X_gt_x <- (length(segment_one[segment_one == TRUE])) / (length(segment_one))

segment_two <- (X > y)
probability_X_gt_y <- (length(segment_two[segment_two == TRUE])) / (length(segment_two))

probability_of_a <- ((probability_X_gt_x * probability_X_gt_y)) / (probability_X_gt_y)
probability_of_a
```

**Answer: $P(X > x | X > y) = 0.5$**


B. $P(X>x, Y>y)$

```{r, probabilityB}
part_one <- (X > x)
probability_X_gt_x <- (length(part_one[part_one == TRUE])) / (length(part_one))

part_two <- (Y > y)
probability_Y_gt_y <- (length(part_two[part_two == TRUE])) / (length(part_two))

probability_of_b <- (probability_X_gt_x) * (probability_Y_gt_y)
probability_of_b
```

**Answer: $P(X>x, Y>y)$ = 0.375**

C. $P(X<x | X>y)$

```{r, probabilityC}

```

