---
title: "DATA 605 Homework 8"
subtitle: "ASSIGNMENT 8"
author: "Stephen Haslett"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(Matrix)
```


#### Exercise 11 - Page 303
A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?


```{r Exercise-11-solution}
# Sum of the bulbs.
number_of_bulbs <- 100

# The mean lifetime of the bulbs is 1000 hours.
bulbs_mean_lifetime <- 1000

# We have 100 bulbs, so to find the lifetime of the first lightbulb (minimum lifetime),
# we divid the mean lifetime by the number of bulbs.
first_bulp_lifespan <- 1000/100

# The first lightbulb will burn out after 10 hours.
first_bulp_lifespan
```

**Answer:** The expected time for the first of these bulbs to burn out is **10 hours**.


#### Exercise 14 - Page 303
Assume that $x_{1}$ and $x_{2}$ are independent random variables, each having an exponential density with parameter $\lambda$. Show that $Z = x_{1} − x_{2}$ has density:
$$f_{Z} = (1/2)\lambda_{e}^-\lambda|z|$$

**Answer:**

$fX_1(x) = fX_2(x) = \begin{cases} \lambda e^{-\lambda x}, & x \ge 0 \\ 0, & x <0 \end{cases}$

$fZ(z) = \int_{-\infty}^{\infty} fX_1(x)fX_2(x - z)dx$

$= \int_{0}^{\infty} \lambda e^{-\lambda x}\lambda e^{-\lambda (x-z)}dx$

$= \int_{0}^{\infty} \lambda^2e^{-2\lambda x + \lambda z}dx$

$= \lambda e^{\lambda z} \int_{0}^{\infty} e^{-2 \lambda x} dx$

$= \frac{1}{2}  \lambda e^{-\lambda |z|}$


#### Exercise 1 - Pages 320 - 321
Let $X$ be a continuous random variable with mean $µ = 10$ and variance $σ^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.

(a) $P(|X − 10| ≥ 2)$.
(b) $P(|X − 10| ≥ 5)$.
(c) $P(|X − 10| ≥ 9)$.
(d) $P(|X − 10| ≥ 20)$.

  
```{r Exercise-1-chebyshev}
# Create a function to calculate Chebyshev’s Inequality.
chebyshev_inequality_calculator <- function(epsilon) {
  variance <- 100 / 3
  probability_x <- variance / epsilon^2

  return(round(probability_x, 4))
}
```

```{r Exercise-1-a-solution}
# Find the upper bound of P(|X − 10| ≥ 2.
answer_a <- chebyshev_inequality_calculator(2)
paste("P(|X − 10| ≥ 2 =", answer_a)
```


```{r Exercise-1-b-solution}
# Find the upper bound of P(|X − 10| ≥ 5).
answer_b <- chebyshev_inequality_calculator(5)
paste("P(|X − 10| ≥ 5) =", answer_b)
```


```{r Exercise-1-c-solution}
# Find the upper bound of P(|X − 10| ≥ 9).
answer_c <- chebyshev_inequality_calculator(9)
paste("P(|X − 10| ≥ 9) =", answer_c)
```


```{r Exercise-1-d-solution}
# Find the upper bound of P(|X − 10| ≥ 20).
answer_d <- chebyshev_inequality_calculator(20)
paste("P(|X − 10| ≥ 20) =", answer_d)
```

**Answers:**

(a) P(|X − 10| ≥ 2 = **8.3333**

(b) P(|X − 10| ≥ 5) = **1.3333**

(c) P(|X − 10| ≥ 9) = **0.4115**

(d) P(|X − 10| ≥ 20) = **0.0833**

