---
title: "DATA 605 Discussion 12"
subtitle: "Is there a relationship between a person's height and their father's height?."
author: "Stephen Haslett"
date: "11/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the necessary libraries.
library(knitr)
```

**Using simple linear regression, we want to answer the question "Is there a relationship between a person's height and their father's height?"**

### The Dataset
For this investigation we will use a dataset compiled by [Francis Galton](https://en.wikipedia.org/wiki/Francis_Galton) that contains data about the heights of children and their parents.

First we will import the dataset into R and take a brief look at the data.
```{r dataSetUp}
# Import Galton's dataset.
family_height_data <-  read.csv("http://cknudson.com/data/Galton.csv")
kable(head(family_height_data), format = 'markdown')

# Count the number of observations in the dataset.
nrow(family_height_data)
```

As we can see from the above, the dataset contains 6 variables: _Family ID_, _Father's Height_, _Mother's Height_, _Gender_, _Height_, and _Number of Kids_, and **898** observations.


### Data Exploration
The first step of our data exploration is to look at a summary of the data.

```{r, dataSummary}
# Summarize the dataset.
options(knitr.kable.NA = '')
kable(summary(family_height_data), format = 'markdown')
```




```{r, medianAssetsScatter}

```




## Regression Model

```{r, dataLinearRegression}

```



```{r, data}
# Ensure R outputs the summary data in a non scientific format so that it is more readable
# (i.e. '-691.837588' rather than '-6.918e+02').
#options(scipen = 4)

# Output a summary of the data.

```


## Residual Analysis

```{r residualAnalysis}

```




