---
title: "DATA 605 Discussion 8"
subtitle: "Exercise 16, Chapter 8, Page 323"
author: "Stephen Haslett"
date: "9/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Exercise 16
Let Z = X/Y where X and Y have normal densities with mean 0 and standard deviation 1. Then it can be shown that Z has a Cauchy density.

#### Part A
(a) Write a program to illustrate this result by plotting a bar graph of 1000 samples obtained by forming the ratio of two standard normal outcomes. Compare your bar graph with the graph of the Cauchy density.


```{r 16ANormalDistribution}
standard_density_graph <- hist(rnorm(1000) / rnorm(1000), breaks = 1000, xlim = c(-20, 20),
                                  xlab = 'Samples', ylab = 'Density', col = 'green',
                                  main = 'Standard Densities')
```

```{r 16ACauchyDistribution}
cauchy_density_graph <- hist(rcauchy(1000), breaks = 1000, xlim = c(-20, 20),
                                  xlab = 'Samples', ylab = 'Density', col = 'red',
                                  main = 'Cauchy Densities')
```

**Answer:**


#### Part B
(b) We have seen that the Law of Large Numbers does not apply to the Cauchy density (see Example 8.8). Simulate a large number of experiments with Cauchy density and compute the average of your results. Do these averages seem to be approaching a limit? If so can you explain why this might be?  