---
title: "DATA 605 Homework 2"
subtitle: "ASSIGNMENT 2"
author: "Stephen Haslett"
date: "9/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

## Problem Set 1

#### (1) Show that $A^{T} A \ne AA^{T}$ in general. (Proof and demonstration.)

If we take matrix $A$ to be of the type $m \times n$, then $A^{T}$ will be of the type $n \times m$ as by definition, $A^{T}$ is a flipped version of $A$ (_i.e row 1 of $A$ becomes column 1 of $A^{T}$, row 2 of $A$ becomes column 2 of $A^{T}$, etc._).

Multiplying $A^{T}$ ($n \times m$) by $A$ ($m \times n$) will result in a square $n \times n$ matrix. If we reverse the multiplication so that it is $A$ ($m \times n$) by $A^{T}$ ($n \times m$), then we will get an $m \times m$ matrix.

Therefore, as per the above logic, if $m \ne n$, the result of $A^{T} A$ ($n \times n$) will always be different to that of $AA^{T}$ ($m \times m$). The resulting matrices will not be equal and so we can assert that **$A^{T} A \ne AA^{T}$**.

This is best demonstrated through the use of an example:

```{r solution1}
# Matrix A.
A <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8), nrow = 2, byrow = TRUE)

# Matrix A transposed.
At <- t(A)

# Display Matrix A.
A

# Display Matrix A transposed.
At

# Multiple A x At. This results in a 2 X 2 matrix.
A %*% At

# Multiple At x A. This results in a 4 x 4 matrix.
At %*% A
```

**Answer:** As per the above example, multiplying $A \times A^{T}$ results in a 2 x 2 matrix, whereas $A^{T} \times A$ results in a 4 x 4 matrix and thus $A^{T} A \ne AA^{T}$.
  
#### (2) For a special type of square matrix $A$, we get $A^{T} A = AA^{T}$. Under what conditions could this be true?

**Answer:** This is true for symmetric matrices where $A = A^{T}$.

Again, this is best demonstrated by an example:

```{r solution2}
# Define Matrix A, and Matrix At.
A <- matrix(c(1, 2, 0, 2, 1, 0, 0, 0, 5), nrow = 3, byrow = TRUE)
At = t(A)

# Display Matrix A.
A

# Display Matrix At.
At

# Confirm that Matrix A is equal to Matrix At.
A == At

# Confirm that AtA = AAt.
(A %*% At) == (At %*% A)
```




## Problem Set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track fights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.

#### Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

The following function factorizes square matrix A into LU.

```{r set2solution}
# Function that factorizes A into LU.
a_to_lu_factorization <- function(A) {
  # Check that passed matrix 'A' is squared. If not, print an error message and return 'NA'.
  if (dim(A)[1] != dim(A)[2]) {
    print('Only squared matrices are valid. Please ensure you are passing the correct matrix type.')

    return(NA)
  }
  else {
    # Otherwise, continue with calculations and return the expected result.
    L <- matrix(c(rep(0, nrow(A) * nrow(A))), nrow = nrow(A))

    # This is not a neccessary variable assignment, but for the sake of readability,
    # assign the value of passed parameter 'A', to a new variable 'U'.
    U <- A

    for (column_n in 1:(nrow(U))) {
      if (column_n <= nrow(U)) {
        L[column_n:nrow(U), column_n] <- U[column_n:nrow(U), column_n] / U[column_n, column_n]
      }

      if (column_n < nrow(U)) {
        for(k in column_n:(nrow(U) -1)) {
          k <- k + 1

          # Create 0 row values.
          U[k, ] <- U[k, ] + (U[column_n, ] * (-1 * U[k, column_n] / U[column_n, column_n])) 
        }
      }
    }

    # Return list of results.
    return(list(L, U))
  }
}
```

```{r set2example}
# Define a sample matrix to pass into our function.
sample_matrix_a <- matrix(c(-3, 1, 2, 6, -5, 2, 9, 5, -6), nrow = 3, byrow = TRUE)

# Run the function.
LU <- a_to_lu_factorization(sample_matrix_a)

# Display the output of the function.
LU
```