---
title: "DATA 605 Discussion 14"
subtitle: "Chapter 7.4, Exercises 7.4, Exercise 5, Page 386"
author: "Stephen Haslett"
date: "12/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pracma)
```

### Assignment Instructions
Using R, provide the solution for any exercise in either Chapter 4 or Chapter 7 of the calculus textbook.


#### Exercise 5
Find the arc length of the function on the given interval:

5. $f(x) = \frac{1}{3}x^{\frac{3}{2}}-x^{\frac{1}{2}} \ on \ [0,1]$.

**Solution**

Define the function in R and plot the curve for a visual reference.
```{r exercise5FunctionAndPlot}
# Define the function using R syntax.
f <- function(x) {
  1/3 * x ^ (3/2) - x ^ (1/2)
}

x <- seq(0, 1, 0.01)
curve(f(x), col = 'blue', ylab='f(x)')
```


Now find the arc length of the function on the given interval using the __pracma__ library's [arclength()](https://www.rdocumentation.org/packages/pracma/versions/1.9.9/topics/arclength) function.

```{r exercise5Answer}
# The arclength() function expects the first paremeter to be parameterized, so we
# need to create a parameterize version of our function before passing it to arclength().
par_function <- function(x) c(x, f(x))
arclength(par_function, a = 0, b = 1)
```


#### Answer
**The arc length of the function on the given interval is**