---
title: "DATA 605 Final Project"
author: "Stephen Haslett"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)

# Disable scientific notation.
options(scipen = 999)
```
\ 

#### Problem 1

Using R, generate a random variable $X$ that has 10,000 random uniform numbers from $1$ to $N$, where $N$ can be any number of your choosing greater than or equal to $6$. Then generate a random variable $Y$ that has $10,000$ random normal numbers with a mean of $\mu =\sigma=(N+1)/2$.  

```{r, randomVariableX}
set.seed(100)

# Set N to any number greater than or equal to 6 (in this case, 8).
N <- 8

# Generate a random variable X that has 10,000 random uniform numbers from 1 to N.
X <- runif(10000, 1, N)

# Generate a random variable Y that has 10,000 random normal numbers with the requested mean.
Y <- rnorm(10000, mean = (N+1)/2, (N+1)/2)
```

\ 

##### _Problem 1 - Probability_

Calculate as a minimum the below probabilities $a$ through $c$.  Assume the small letter "$x$" is estimated as the median of the $X$ variable, and the small letter "$y$" is estimated as the 1st quartile of the $Y$ variable. Interpret the meaning of all probabilities.

```{r, probability}
# Small x is estimated as the median of the X variable.
x <- round(median(X), 2)
x

# Small y is estimated as the 1st quartile of the Y variable.
y <- round(quantile(Y, 0.25), 2)
y
```

**Answer**: Small x is equal to **4.48** (median of the X variable), small y is equal to **1.45** (1st quartile of the Y variable). 


A. $P(X>x \ | \ X>y)$

**Interpretation**:

_Probability that X is greater than its median given that X is greater than the first quartile of Y_.

$$P(X>x \ | \ X>y) = \frac{P(X>x \ , \ X>y)}{P(X>y)}$$

```{r, probabilityA}
# Define the events.
event_one <- (X > x)
event_two <- (X > y)

# P(X>x and X>y).
a_and_b <- length(X[event_one & event_two]) / length(X)

# P(X>y).
b <- length(X[event_two]) / length(X)

# P(X>x | X>y).
probability <- a_and_b / b

answer <- round(probability, 2)
answer
```

**Answer: $P(X > x \ | \ X > y)$ = 0.53**

\ 

B. $P(X>x, Y>y)$

**Interpretation**:

_Probability that $X$ is greater than $x$, and $Y$ is greater than $y$_. 

```{r, probabilityB}
# Define the events.
event_one <- (X > x)
event_two <- (Y > y)

# P(X > x).
X_gt_x <- length(X[event_one]) / length(X)

# P(Y > y).
Y_gt_y <- length(Y[event_two]) / length(Y)

probability  <- X_gt_x * Y_gt_y

answer <- round(probability, 2)
answer
```

**Answer: $P(X>x, Y>y)$ = 0.37**

\ 

C. $P(X<x \ | \ X>y)$

**Interpretation**:

_Probability that $X$ is less than its median given that it is greater than the first quantile of $Y$_.

```{r, probabilityC}
# Define the events.
event_one <- (X < x)
event_two <- (X > y)

# P(X > x & X > y).
a_and_b  <- length(X[event_one & event_two]) / length(X)

# P(X > y).
b <- length(X[event_two]) / length(X)

probability <- a_and_b / b 

answer <- round(probability, 2)
answer
```

**Answer: $P(X<x \ | \ X>y)$ = 0.47**

\ 

Investigate whether $P(X>x \ and \ Y>y) = P(X>x)P(Y>y)$ by building a table and evaluating the marginal and joint probabilities.

```{r, probabilityTable}


part_one <- (X > x)
prob_X_gt_x <- (length(part_one[part_one == TRUE])) / (length(part_one))

part_two <- (Y > y)
prob_Y_gt_y <- (length(part_two[part_two == TRUE])) / (length(part_two))

results_table <- data.table(
  Event = c('(X>x)', '(Y>y)', '(X>x)*(Y>y)', '(X>x and Y>y)'),
  Xx = c(prob_X_gt_x, prob_Y_gt_y, prob_X_gt_x * prob_Y_gt_y, prob_X_gt_x * prob_Y_gt_y),
  Yy = c(prob_Y_gt_y, prob_X_gt_x, prob_X_gt_x * prob_Y_gt_y, prob_X_gt_x * prob_Y_gt_y)
)

results_table
```

\ 

#### Problem 2

Register for Kaggle.com and compete in the House Prices: Advanced Regression Techniques competition.

**Competition Objective**

Ask a home buyer to describe their dream house, and they probably won't begin with the height of the basement ceiling or the proximity to an east-west railroad. But this playground competition's dataset proves that much more influences price negotiations than the number of bedrooms or a white-picket fence.

With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this competition challenges you to predict the final price of each home.

\ 

**Import the training and test datasets and summarize them**.

```{r, loadData}
# Pull in the train and test datasets.
training_dataset <- read.csv('https://raw.githubusercontent.com/stephen-haslett/data605/data605-final-exam/train.csv')
test_dataset <- read.csv('https://raw.githubusercontent.com/stephen-haslett/data605/data605-final-exam/test.csv')
```

\ 

**Snapshot of the training dataset**.

```{r, trainingSnapshot}
head(training_dataset, 1)
```


**Summarize the training dataset**.

```{r, trainingSummary}
summary(training_dataset)
```

\ 

**Snapshot of the test dataset**.

```{r, testSnapshot}
head(test_dataset, 1)
```

**Summarize the test dataset**.

```{r, testSummary}
summary(test_dataset)
```

\ 

##### Descriptive and Inferential Statistics

**1. Provide univariate descriptive statistics and appropriate plots for the training data set**.

```{r, salesPrice}
# Summarize the training dataset's SalePrice variable.
summary(training_dataset$SalePrice)
```

\ 

**Create a histogram of sales prices**.

```{r, salesPricePlot}
hist(training_dataset$SalePrice,
     xlab = 'Sale Price',
     main = 'Distribution of Sales Prices',
     col = 'darkgreen')
```

\ 

**Create a QQ plot of sales prices**.

```{r, salesPriceQQPlot}
qqnorm(training_dataset$SalePrice, col = 'darkred')
qqline(training_dataset$SalePrice)
```

\ 

**2. Provide a scatterplot matrix for at least two of the independent variables and the dependent variable**.

```{r, scatterplotMatrix}
# Define the variables to include in the matrix.
sale_price <- training_dataset$SalePrice
lot_area <- training_dataset$LotArea
gr_liv_area <- training_dataset$GrLivArea
garage_area <- training_dataset$GarageArea

# Plot the matrix.
plot_data <- data.frame(sale_price, lot_area, gr_liv_area, garage_area)
pairs(plot_data, main = 'Scatterplot Matrix', col = '#50394c')
```